\documentclass{article}
\usepackage[margin=1in]{geometry}
\setlength{\parindent}{0em}
\setlength{\parskip}{0.5em}

\uspackage{enumitem}
\setlist[description]{%
    \ttfamily\textcolor{cadet}{#1}
}

\begin{document}

\section{Unsorted stuff}
[gg] Move to first line
[''/``] Return to line/position of cursor\ldots?
[:vsplit file]
[:split file] Same as :new file?
[^] Go to beginning of text in line
[_] Go to beginning of text in line
[e/E, b/B] puctuation/no punc
[ctrl u/b, d/f]
[yw], [w], [W]
[:#1], [:#2] ... ?
[u]
[:u] undo last COMMAND
[:e!] Read original file back
[U] in visual mode, changes to all caps, letters that are already caps
    will not be changed.
[ctrl ^] Open last documnet (e.g. do  :!e new, then :CTRL ^)???
[^T] set autoindent tab one tabstop to the right
[^D] ... left
[#|<CR>] Go to column #
[:retab] replace existing tabs with spaces

Markdowns: zo, zc, zR

Mappings: vmap - visual, imap - insert, nmap - normal

search and replace:
s/find/replace/g vs. %s/find/replace/g
:s/\*/g   ... must escape special characters

Apply things to multiple lines :l1,l1<command>
:l1,l2d
:l1,l2y
:l1,l2s/find/replace/g
Or apply command to visually selected lines

highlight, than '=' to properly space things

in visual mode, 'gq' to change lines to textwidth width.

:qa  Quit all (buffers, etc.)

Buffers?

Find:
fx<CR>  find character after cursor
Fx<CR>  find character before cursor

:SyntasticToggleMode    active <--> passive
:SyntasticCheck ctermbg = #2f0000
highlight Syntastic error?

:help <expression>

\section{Cursor movement}

\begin{tabular}{l l}
    \textbf{command} & \textbf{description}\\
    \hline
    \verb|z<CR>| & makes current line the top line of the page\\
    \verb|z.| & makes current line the middle line of the page\\
    \verb|z-| & move current line to bottom of page\\
    \verb|%| & move to matching bracket, parentheses, etc.\\
\end{tabular}


\section{Searching}

\section{Mapping}
\begin{minipage}[t]{0.4\textwidth}
    \texttt{:map <ENTER> a<ENTER><ESC}\\
    \texttt{:map <F9> 070l}\\
\end{minipage}
\begin{minipage}[t]{0.7\textwidth}
    use return to break lines without going into insert mode.\\
    shortcut for moving to right edge of screen (not working).\\
\end{minipage}

\section{Misc}

\subsection{``set'' options}
\subsubsection{Syntax}
\texttt{:set \emph{mycommand}}
\subsubsection{Examples}
\par\texttt{:set incsearch} to move to matching pattern when searching.
\par\texttt{:set ignorecase}
\par\texttt{:set tabstop = 4} tab four spaces instead of the default of eight
\par\texttt{:set expandtab} tabs are converted to spaces (useful since not every
computer handles tabs the same way, but note that makefiles will not
work without actual tabs).
\par\texttt{set spell} Spellcheck! Do this in command mode; leaving it on
will probably get really annoying.

\section{Buffers}

\subsection{Other}
\par\texttt{:edit!} or \texttt{:e!} Undoes all edits since last write
\par\texttt{:cal cursor(row, column)} Go to row, column. What can you put
in for ``row'' to indicate the current line the cursor is in?
\par\texttt{:help map-which-keys} Tells what keys are currenly in use.


\end{document}
